---
title: "data_wrangling_i"
author: "Ani Bilazarian"
date: "9/17/19"
output: github_document
---


```{r}
library(tidyverse)
litters_data = read_csv(file = "./data/FAS_litters.csv")
litters_data=janitor:: clean_names(litters_data)
litters_data
```

```{r}
view(litters_data)
```

```{r setup, include=FALSE}
knitr:: opts_chunk$set(echo=TRUE)
```

# Load in the pups data 
```{r}
pups_data = read_csv(file= "./data/FAS_pups.csv")
pups_data=janitor:: clean_names(pups_data)
pups_data
```

# Absolute Path 
Start from the beginning and give step by step directions to where you're going -person specific-
ex. "/Users/jeffgoldsmith/Dropbox/Work/Teaching/P8105/fall_2019/p8105.github.io/data/FAS_litters.csv"


# Relative Path 
relative from a common starting point, so useful for other people 
ex. getwd()  [1] "/Users/jeffgoldsmith/Dropbox/Work/Teaching/P8105/fall_2019/p8105.github.io"


# CLEAN NAME FUNCTION
you are using this in the "janitor" package, you can also put in library(janitor) so your code file has it and you don't have to put janitor:: clean names each time 
but the janitor package may have naming conflicts, so keeping janitor:: Clean names allows you to JUST use the clean names function

# SKIM FUNCTION 
after reading in a new data set, the skim function allows you to get a high level look of the data " skimr::skim(litters_data)" 
-- it is within the skimr package 

# VIEWING DATA & CLEANING TIPS 
if you want to view data do it in the CONSOLE not on the RMD file... makes r studio go crazy! 

* "col_names:" usually TRUE. If FALSE, column names are X1, X1, … . You can also supply column names.

* "na:" string vector containing character expressions for missing values.

* "skip:" number of rows to skip before reading data.

ex. litters_data = read_csv(file = "./data/FAS_litters.csv",
  skip = 10, col_names = FALSE)
  
# PLAY WITH COLUMN TYPES 
This doesn't always happen when you import data, but if you want to be very careful about how your data is being read in, this is helpful 
```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(
    Group = col_character(),
    `Litter Number` = col_character(),
    `GD0 weight` = col_double(),
    `GD18 weight` = col_double(),
    `GD of Birth` = col_integer(),
    `Pups born alive` = col_integer(),
    `Pups dead @ birth` = col_integer(),
    `Pups survive` = col_integer()
  ))
```

tail(litters_data)  


